---
title: "Covid91 vaccine study Final2024"
author: "Lacei Brown"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "Summer 2025" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
Males = filter(FinalData, sex =="M")
Female = filter(FinalData, sex =="F")
gay = filter(FinalData, LGBTQ =="gay")
Druguser = filter(FinalData, DrugUser =="yes")
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

##Barplot
```{r}
dd2 <- Males %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/ sum(count))
# the group_by followed by summarize (count=n ())

basicC <- ggplot(dd2,aes(x=infected,y=count, fill=treatment)
)

basicC + geom_bar (stat="identity", position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment))
basicCC + geom_bar(stat="identity", position = "dodge")
```

The male popultion and their connection with COVID-19 infection through vaccination or placebo are shown in this plot. 48 according to the graph about 25% of those who took the vaccine actually got COVID-19, so about roughly 75% of those who received a placebo did. COmpared to individuals who received the vaccine alone, almost half of those who took the placebo developed COVID-19. It seems that the vaccine is a good way to guard against covid.

## Numerical Results

```{r}
table1 <- xtabs(~infected + treatmeant, data=Males)
rowPerc(table1)
colPerc(table1)
```


The percentages if males who contracted COVID-19 after receiving the vaccine and a placebo are displayed ib these numerical results. About 25% of the males who received the vaccine developed COVID-19, whereas 75% of males who recieved the placebo developed COVID-19. The vaccination is unquestionably successful in protecting males against Covid-19.

## Inferential

```{r}
chisq.test(table1)
chisqtestGC(table1)
Malesfish <- fisher.test(table1)
Malesfish
```

The p-valuse displayed abaove is at 2.2e-16. This number is acceptable. This number shows that these outcomes are probably the result of chance.


```{r}
effect <- Malesfish$p.value/(Malesfish$p.value +1)
effectivenessMales = Malesfish$estimate/(Malesfish$estimate +1)*100
print(effectivenessMales)
```

odds ratio 73.79183

## Females


## LGBTQ



## Druggies


# Overall Results and Conclusions