---
title: "Covid91 vaccine study Final2024"
author: "Lacei Brown"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "Summer 2025" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
Males = filter(FinalData, sex =="M")
Female = filter(FinalData, sex =="F")
gay = filter(FinalData, LGBTQ =="gay")
Druguser = filter(FinalData, DrugUser =="yes")
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

##Barplot
```{r}
dd2 <- Males %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/ sum(count))
# the group_by followed by summarize (count=n ())

basicC <- ggplot(dd2,aes(x=infected,y=count, fill=treatment)
)

basicC + geom_bar (stat="identity", position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment))
basicCC + geom_bar(stat="identity", position = "dodge")
```

The male popultion and their connection with COVID-19 infection through vaccination or placebo are shown in this plot. 48 according to the graph about 25% of those who took the vaccine actually got COVID-19, so about roughly 75% of those who received a placebo did. COmpared to individuals who received the vaccine alone, almost half of those who took the placebo developed COVID-19. It seems that the vaccine is a good way to guard against covid.

## Numerical Results

```{r}
table1 <- xtabs(~infected + treatmeant, data=Males)
rowPerc(table1)
colPerc(table1)
```


The percentages if males who contracted COVID-19 after receiving the vaccine and a placebo are displayed ib these numerical results. About 25% of the males who received the vaccine developed COVID-19, whereas 75% of males who recieved the placebo developed COVID-19. The vaccination is unquestionably successful in protecting males against Covid-19.

## Inferential

```{r}
chisq.test(table1)
chisqtestGC(table1)
Malesfish <- fisher.test(table1)
Malesfish
```

The p-valuse displayed abaove is at 2.2e-16. This number is acceptable. This number shows that these outcomes are probably the result of chance.


```{r}
effect <- Malesfish$p.value/(Malesfish$p.value +1)
effectivenessMales = Malesfish$estimate/(Malesfish$estimate +1)*100
print(effectivenessMales)
```

odds ratio 73.79183

```{r}
low=(Malesfish$conf.int[1]/(Malesfish$conf.int[1] +1))*100
hi=(Malesfish$conf.int[2]/(Malesfish$conf.int[2] +1))*100
options(digits=2)
```

## Effectiveness Confidence Interval `r low`  to `r hi`

The success rate of the COVID-19 vaccine for men is seen above. The vaccine is considered successful if its efficacy is greater that 50%. We may assume from the above time period that the vaccine effectively prevents males from contracting covid 19. The odds ratio provides an estimate of the vaccine's effectiveness. As can be seen above, the men odds ration is 70. For the men, the efficacy confidence interval is between 68.85 and 78.32


## Females

##Barplot

```{r}
dd2 <- Female %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/ sum(count))
# the group_by followed by summarize (count=n ())

basicC <- ggplot(dd2,aes(x=infected,y=count, fill=treatment)
)

basicC + geom_bar (stat="identity", position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment))
basicCC + geom_bar(stat="identity", position = "dodge")
```

The female population and its association with COVID-19 infection through vaccination or placebo are shown in this plot. According to the plot above, about 35% of those who received the vaccine actually got OVID-19, where roughly 65% of those who received a placebo did. Compared to individuals who received the vaccine alone, a significantly greater proportion of those who took that placebo ddeveloped COVID-19. It seems that the vaccine is a good way to guard against COVID-19.

## Numerical Results

```{r}
table2  <- xtabs(~infected + treatmeant, data=Males)
rowPerc(table2)
colPerc(table2)
```

The percentages of females who contracted COVID-19 after receiving the vaccine and a placebo are displayed in these numerical results. About 35% of the men who receieved the vaccine developed COVID-19, whereas 65% of the women who received the placebo did so. The vaccine is advised since it is evidently effecient in preventing conid-19 in males. The risk of gettung covid-19 was higher for those who took the placebo.

## Inferential

```{r}
chisq.test(table1)
chisqtestGC(table1)
Femalefish <- fisher.test(table1)
Femalefish
```  

The p-value displayed above is at 2.2e-16. This is an okay number. Seeing this number means that these results are likelt to happen by chance.

```{r}
effect <- galsfish$p.value/(galsfish$p.value +1)
effectivenessgal = galsfish$estimate/(galsfish$estimate +1)*100
print(effectivenessgal)
```

```{r}
low=(galsfish$conf.int[1]/(galsfish$conf.int[1] +1))*100
hi=(galsfish$conf.int[2]/(galsfish$conf.int[2] +1))*100
options(digits=2)
```

## Effectiveness Confidence Interval `r low`  to `r hi`

The COVID-19 effectiveness for girls (females) is shown above. The vaccine is considered successful if its efficacy is greater than 50%. We can conclude that the vaccine is very efficient at preventing COVID-19 in girls based on the interval shown above. 60% or above is a sign of an effective vaccine. The effectiveness of this vaccine in preventing women from contracting COVID-19 is excellent. The odds ratio provides an estimate of the vaccine's effectiveness. As can be seen above, the gals' odds ratio is 67. The females' efficacy confidence interval is from 63.15 to 70.99.

## LGBTQ

### Barplot

```{r}
dd2 <- gay %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This barplot shows the percentage of people who got covid-19 versus those who stayed healthy, based on whether they received the vaccine or a placebo, within a population that identifies as gay. The red bars represent the placebo group, and the blue bars represent the vaccine group. A much higher percentage of the placebo group got covid-19 compared to the vaccine group, showing that the vaccine was effective in reducing infections. This is the second of two plots. Unlike the first plot (which shows raw counts), this one uses percentages to make it easier to compare between groups.

### Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data=gay)
rowPerc(table1)
colPerc(table1)
```

This shows that the vaccine significantly reduced the risk of infection among gay participants. It was highly effective compared to the placebo. The odds of getting covid-19 with the placebo were much higher, meaning the vaccine helped protect against infection — not the opposite. Therefore, the data supports the use of the vaccine in this group.

### Inferential 

```{r}
chisq.test(table1)
chisqtestGC(table1)
gaysfish <- fisher.test(table1)
gaysfish
```

The p-value from the Chi-square test is 6e-06, which is extremely small. This means that the differences in infection rates between the vaccine and placebo groups are statistically significant and unlikely to be due to chance alone. The Fisher’s Exact Test reports an odds ratio of 0.4. This means that gay individuals who received the vaccine were 60% less likely to get covid-19 compared to those who received the placebo. Because the p-value is so small, we reject the null hypothesis and accept the alternative: there is a relationship between receiving the vaccine and reduced covid-19 infection.

```{r}
effect <- gaysfish$p.value/(gaysfish$p.value +1)
effectivenessgays = gaysfish$estimate/(gaysfish$estimate +1)*100
print(effectivenessgays)
```

```{r}
low=(gaysfish$conf.int[1]/(gaysfish$conf.int[1] +1))*100
hi=(gaysfish$conf.int[2]/(gaysfish$conf.int[2] +1))*100
options(digits=2)
```

## Effectiveness Confidence Interval `r low`  to `r hi`

This shows how well the covid-19 vaccine worked for people who identify as gay. The confidence interval is from 20.51% to 37.97%, which means the vaccine helped reduce the chance of getting covid-19, but not as much as hoped. The odds ratio is 0.29, meaning those who got the vaccine were less likely to get covid-19 than those who got the placebo.
So even though the vaccine wasn’t super strong in this group, it still offered protection.

## Druggies

### Barplot

```{r}
dd2 <- Druguser %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This barplot shows the percentage of people who contracted COVID-19 versus those who remained uninfected, based on whether they received the vaccine or a placebo, within a population of individuals who reported drug use. The red bars represent the placebo group, and the blue bars represent the vaccine group. A significantly higher percentage of the placebo group contracted COVID-19 compared to the vaccine group, indicating that the vaccine was effective in reducing infections among drug users. This is the second of two plots. Unlike the first plot, which displays raw counts, this plot uses percentages to better compare infection rates between the treatment groups.

### Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data= Druguser)
rowPerc(table1)
colPerc(table1)
```

These numbers clearly demonstrate that the vaccine is effective in protecting drug users against COVID-19, making vaccination strongly recommended for this group. The odds of getting COVID-19 while vaccinated are roughly 1:2, meaning drug users who received the vaccine were about half as likely to contract COVID-19 compared to those who took the placebo.

### Inferential 


```{r}
chisq.test(table1)
chisqtestGC(table1)
drugfish <- fisher.test(table1)
drugfish
```

The chi-square test produces a p-value of 0.002, which is very small. This indicates that the observed association between treatment (vaccine vs. placebo) and COVID-19 infection among drug users is unlikely to have occurred by chance. Fisher’s Exact Test estimates an odds ratio of approximately 1.8, meaning drug users who received the vaccine were about half as likely to contract COVID-19 compared to those who took the placebo. Based on these results, we reject the null hypothesis of no association and accept the alternative hypothesis: the vaccine is effective in reducing the risk of COVID-19 infection among drug users.

```{r}
effect <- drugfish$p.value/(drugfish$p.value +1)
effectivenessdrug = drugfish$estimate/(drugfish$estimate +1)*100
print(effectivenessdrug)
```

```{r}
low=(drugfish$conf.int[1]/(drugfish$conf.int[1] +1))*100
hi=(drugfish$conf.int[2]/(drugfish$conf.int[2] +1))*100
options(digits=2)
```




# Overall Results and Conclusions